
\pagenumbering{Roman}
\setcounter{page}{2}

\normalsize

\chapter*{Declaration of Authorship}
\addcontentsline{toc}{chapter}{Declaration of Authorship}

„I declare in lieu of oath that this thesis is entirely my own work except where
otherwise indicated. The presence of quoted or paraphrased material has been clearly
signaled and all sources have been referred. The thesis has not been submitted for a degree
at any other institution and has not been published yet.”

\chapter*{Preface}
 \addcontentsline{toc}{chapter}{Preface}
%
%Eloquent expression of my gratitude towards everyone involved in this thesis.

How often do you get the chance to say, that for one short moment in time, you were the best at something? 
\textcite[93]{Heal10} states, that the drop test rig built by WASM is the most sophisticated and best instrumented rig, in the world today. As it is his Alma Mater, he added a modest "perhaps". 
Reading those lines motivated me beyond believe, I knew we had the potential to be better and I knew I wanted to be better. This is why I have to thank Daniel Heal first and foremost.

Secondly I have to thank Biruk Woldhemedin who had to endure the bulk of my misguided enthusiasm and crazy ideas and who kept his humor and his patience even as I cleaned his whiteboard or made fun of his palm tree and helped me navigate the treacherous rapids of Swedish culture. Keep on being awesome.

I want to thank my supervisors at Montanuniversity Leoben, Horst Wagner and Tobias Ladinig for supporting me and not giving up on me, even when I acted too laissez-faire for their taste or sent them too many too long emails.

Matthias Wimmer, I must thank you for making this possible and supporting me with everything. 

I want to thank SODA, the board game club of Kiruna, and all of its members for helping me keep my sanity through the long dark polar night and the long bright polar day. My role-play friends: Adam, Jule, Robert, Tony, William and Flimbimsy. I hope you enjoyed roaming the galaxy and victorian London as much as I did.
Especially Ingrid, thanks for your support with my real and fictional relationship troubles. I felt more at home in your apartment than anywhere I lived in the past few years. 

I want to thank my family and friends for supporting me. My sister who found time for me next to her two studies, two cats and everything else. My friends in Leoben who kept in contact even though I was so far away.
Erik, thank you for enduring long hours of talking about our feelings, great recommendations in music and movies and your patience.
Gü, you helped me a lot with your spreadsheets, your interesting ideas and your patience with my silliness. 

With this, all that is left for me to say is: good luck to the next person who gets to work with this rig. I hope this thesis, my notes, my code and the decisions I made help you more than they hinder you and I look forward to reading what you manage to do with this impressive machine. Good luck.